---
title: "Figure-3"
author: "Mac Campbell"
date: "2023-02-13"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message=FALSE)
```


```{r}
library(tidyverse)
library(ggtree)
library(ape)
library(viridis)
```

## Phylogeny

Generating a data set for concated and species tree analyses. Will reduce sampling of non-wow clade fish.

These our are downsampled and complete read counts.

```{r}
b3b<-read_tsv("meta/b3b.tsv")
```

### Mountains
Get 2-4 or each location of Pantosteus

```{r}
mountains<-read_tsv("outputs/300/mymountains.tsv")
tahoe<-read_tsv("outputs/300/mytahoe.tsv")
```

```{r}
mpc<-b3b %>% dplyr::select(`Sample ID`, `Species Common Name`, Location, Counts, NewPath) %>% filter(`Sample ID` %in% mountains$sample) %>% group_by(Location) %>% arrange(-Counts) %>%
  filter(Location=="Poore Creek, CA") %>% top_n(n=2, wt=Counts)

mtr<-b3b %>% dplyr::select(`Sample ID`, `Species Common Name`, Location, Counts, NewPath) %>% filter(`Sample ID` %in% mountains$sample) %>% group_by(Location) %>% arrange(-Counts) %>%
  filter(Location=="Truckee River, NV") %>% top_n(n=2, wt=Counts)
```

```{r}
og<-bind_rows(mpc,mtr)
```


### Tahoes
```{r}
tpc<-b3b %>% dplyr::select(`Sample ID`, `Species Common Name`, Location, Counts, NewPath) %>% filter(`Sample ID` %in% tahoe$sample) %>% group_by(Location) %>% arrange(-Counts) %>%
  filter(Location=="Poore Creek, CA") %>% top_n(n=2, wt=Counts)

ttr<-b3b %>% dplyr::select(`Sample ID`, `Species Common Name`, Location, Counts, NewPath) %>% filter(`Sample ID` %in% tahoe$sample) %>% group_by(Location) %>% arrange(-Counts) %>%
  filter(Location=="Truckee River, NV") %>% top_n(n=2, wt=Counts)

tcr<-b3b %>% dplyr::select(`Sample ID`, `Species Common Name`, Location, Counts, NewPath) %>% filter(`Sample ID` %in% tahoe$sample) %>% group_by(Location) %>% arrange(-Counts) %>%
  filter(Location=="East Fork Carson River, NV") %>% top_n(n=2, wt=Counts)

```

```{r}
tahoes<-bind_rows(tpc,ttr,tcr)
```


### Sacs

```{r}
additional<-read_csv("meta/additional-meta.csv")

lac<- b3b %>% left_join(additional, by=c("Sample ID"="GVL Code")) %>% dplyr::select(`Sample ID`, `Species Common Name`, Location, Counts, NewPath, Subspecies) %>% filter(Subspecies=="lacuanserinus") %>%
  arrange(-Counts)  %>% top_n(n=2, wt=Counts) %>% dplyr::select(-Subspecies)

# lac from Ash Creek has generally good coverage, Goose lake dent has 156708 reads (1208-008)

occ<-b3b %>% left_join(additional, by=c("Sample ID"="GVL Code")) %>% dplyr::select(`Sample ID`, `Species Common Name`, Location, Counts, NewPath, Subspecies) %>% filter(Subspecies=="occidentalis") %>%
  arrange(-Counts)  %>% top_n(n=2, wt=Counts) %>% dplyr::select(-Subspecies)

# occidentalis also has low reads.... 
```

```{r}
sacs<-bind_rows(lac, occ)
```

### Modocs
```{r}
modocs<-b3b %>% left_join(additional, by=c("Sample ID"="GVL Code")) %>% dplyr::select(`Sample ID`, `Species Common Name`, Location, Counts, NewPath, Coordinates) %>%
  filter(`Species Common Name`=="Modoc Sucker") %>% 
  arrange(-Counts) %>% top_n(4, Counts) %>% dplyr::select(-Coordinates)

#Two turner creeks and two ash creeks
```


### Klams

Get a couple of each species
```{r}
larges<-b3b %>% dplyr::select(`Sample ID`, `Species Common Name`, Location, Counts, NewPath) %>% filter(`Species Common Name`=="Klamath Largescale Sucker") %>%
  arrange(-Counts)  %>% top_n(n=2, wt=Counts)

smalls<-b3b %>% dplyr::select(`Sample ID`, `Species Common Name`, Location, Counts, NewPath) %>% filter(`Species Common Name`=="Klamath Smallscale Sucker") %>%
  arrange(-Counts)  %>% top_n(n=2, wt=Counts)

losts<-b3b %>% dplyr::select(`Sample ID`, `Species Common Name`, Location, Counts, NewPath) %>% filter(`Species Common Name`=="Lost River Sucker") %>%
  arrange(-Counts)  %>% top_n(n=2, wt=Counts)

shorts<-b3b %>% dplyr::select(`Sample ID`, `Species Common Name`, Location, Counts, NewPath) %>% filter(`Species Common Name`=="Shortnose Sucker") %>%
  arrange(-Counts)  %>% top_n(n=2, wt=Counts)
```

```{r}
klams<-bind_rows(larges,smalls,losts,shorts)
```


### Wow Clade
```{r}
wowsers<-b3b %>% dplyr::select(`Sample ID`, `Species Common Name`, Location, Counts, NewPath) %>% filter(`Species Common Name` %in% c("Wall Canyon Sucker","Warner Sucker","Owens Sucker")) %>%
  arrange(-Counts) 
  
ggplot(wowsers) +
  geom_histogram(aes(x = Counts, fill=`Species Common Name`)) +
  theme_bw()
```

```{r}
wowsers %>% filter(Counts>5e5) %>% group_by(`Species Common Name`,Location) %>% summarize(Count=n())
```

```{r}
keepers<-wowsers %>% filter(Counts>4e5)
```


## Combine

```{r}
myfish<-bind_rows(og,tahoes,sacs,modocs,klams,keepers)
write_tsv(myfish,"outputs/802/myfish.tsv")
myfish %>% group_by(`Species Common Name`) %>% summarize(Count=n())
```


```{r}
nrow(myfish)
write_csv(myfish, "meta/phylo119.csv")
```


```{r}
write_tsv(dplyr::select(myfish %>% ungroup, NewPath), col_names = FALSE, "bamlists/phylo119.bamlist")

write_tsv(dplyr::select(myfish %>% ungroup, `Sample ID`), col_names = FALSE, "outputs/204/samples119.txt")

```
```{sh, eval=FALSE}
srun -p bigmemh -t 18:00:00 --mem=16G --nodes=1 --ntasks=1 angsd -P 6 \
-bam bamlists/phylo119.bamlist \
-out /home/maccamp/wall-canyon/outputs/204/plink \
-anc $HOME/genomes/xyrauchen/Xyrauchen_texanus.faa \
-minInd 113 -minMaf 0.05  -minMapQ 10 -minQ 20 -GL 1 -doMajorMinor 1 -doMaf 1 -SNP_pval 1e-6 \
-doGeno 4 -doPost 1 -postCutoff 0.95 -doPlink 2 >outputs/204/std.out 2>outputs/204/std.err &
```

## Process


```{sh, eval=FALSE}
plink --tped plink.tped --tfam plink.tfam  --out binary --recode --allow-extra-chr --noweb
plink --ped binary.ped --map binary.map --recode vcf --allow-extra-chr -out recode
bcftools +prune -l 0.9 -w 100000 recode.vcf  -Ov -o recode.prune.vcf
bcftools reheader --samples samples119.txt -o recode.prune.reheadered.vcf recode.prune.vcf

```

Convert to phylip, then nexus. Create a missing data set

```{sh, eval=FALSE}
source activate py2; ~/github/mccloud-rrt/vcf2phylip.py -i recode.prune.reheadered.vcf; conda deactivate;
seqConverter.pl -drecode.prune.reheadered.min4.phy -on

#Corrected for ascertainment bias
~/github/mccloud-rrt/103-remove-invariant.py -p recode.prune.reheadered.min4.phy -o recode.prune.reheadered.min4.asc.phy
seqConverter.pl -drecode.prune.reheadered.min4.asc.phy -on

#Now remove missing??
 ~/github/relict-dace/201.1-removeMissing.pl recode.prune.reheadered.min4.asc.phy  > filtered.phy
seqConverter.pl -dfiltered.phy -on
```
    
    
##Concat     

```{sh, eval=FALSE}
iqtree -s recode.prune.reheadered.min4.asc.phy -st DNA -m MFP+ASC -bb 1000 -alrt 1000 -redo
```

Moved to farm:

```{sh, eval=FALSE}
srun -p high -t 05:00:00 iqtree -nt AUTO -s recode.prune.reheadered.min4.asc.phy -st DNA -m MFP+ASC -bb 1000 -alrt 1000 -redo
```


Alignment has 119 sequences with 8248 columns, 8247 distinct patterns
7433 parsimony-informative, 815 singleton sites, 0 constant sites


```{r}
iq<-read.tree("outputs/204/recode.prune.reheadered.min4.asc.phy.contree")
iq<-root(iq, c("MS_PC_01","MS_PC_06","S_TR_17","S_TR_01"))
t<-as.polytomy(iq, feature='node.label', fun=function(x) as.numeric(x) < 75)
t<-ggtree(t)

t$data$Bootstrap<-as.numeric(t$data$label)

d <- t$data
d <- d[!d$isTip,]
d$label <- as.numeric(d$label)
d <- d[d$label >= 90,]

e <- t$data
e <- e[!e$isTip,]
e$label <- as.numeric(e$label)
e <- e[e$label < 90,]
e <- e[e$label >= 75,]

```

```{r}
t %<+% myfish +
  geom_point(data=e,  color="gray50", cex=5, alpha=0.75, pch=18) +
  geom_point(data=d,  color="black", cex=5, alpha=0.75, pch=18) +
#  geom_tippoint(aes(fill=`Species Common Name`,x=x+0.03), pch=21) +
  geom_tiplab(aes(label=Location, color=`Species Common Name`), size=2, align = TRUE) +
#  geom_nodelab() +
  xlim(0,.5)

ggsave("outputs/204/concat-snps.pdf", width=7, height=8)
```






These samples have a high propoirtion of missing data     
(base) Macs-MBP-2:204 mac$ cat dumped.txt | cut -f 1      
MS_PC_03   
MS_PC_02    
934-003    
934-002   
1208-059    
1208-064   
WS_TMC15   
WS_DC01   
WS_TMC02   
WS_TMC05   
WS_TMC12   
WS_DC14  

```{sh, eval=FALSE}
iqtree -s filtered.phy -st DNA -m MFP+ASC -bb 1000 -alrt 1000 -redo
iqtree -s filtered.phy.varsites.phy -st DNA -m MFP+ASC -bb 1000 -alrt 1000 -redo
```


Alignment has 107 sequences with 8168 columns, 8066 distinct patterns

7330 parsimony-informative, 838 singleton sites, 0 constant sites

```{sh, eval=FALSE}
srun -p high -t 05:00:00 iqtree -nt AUTO -s filtered.phy.varsites.phy -st DNA -m MFP+ASC -bb 1000 -alrt 1000 -redo
```

```{r}
iq<-read.tree("outputs/204/filtered.phy.varsites.phy.contree")
iq<-root(iq, c("MS_PC_01","MS_PC_06","S_TR_17","S_TR_01","S_TR_09","S_TR_10","TS_EFCR_13","TS_EFCR_15"))
t<-as.polytomy(iq, feature='node.label', fun=function(x) as.numeric(x) < 75)
t<-ggtree(t) %>% ggtree::rotate(127) %>% ggtree::rotate(146)


t$data$Bootstrap<-as.numeric(t$data$label)

d <- t$data
d <- d[!d$isTip,]
d$label <- as.numeric(d$label)
d <- d[d$label >= 90,]

e <- t$data
e <- e[!e$isTip,]
e$label <- as.numeric(e$label)
e <- e[e$label < 90,]
e <- e[e$label >= 75,]

```

```{r}
t2<-t %<+% left_join(myfish, additional, by=c("Sample ID"="GVL Code")) 
```

```{r}
t2$data$Location<-gsub(", Bishop, CA|  Bishop, CA|, Bishop,CA|,  Bishop, CA","",t2$data$Location)
t2$data$Location<-gsub(", NV|, OR","",t2$data$Location)
```

```{r}
t2 +
  geom_point(data=e,  fill="grey", cex=2, alpha=0.75, pch=21) +
  geom_point(data=d,  fill="black", cex=2, alpha=0.75, pch=21) +
#  geom_tippoint(aes(fill=`Species Common Name`,x=x+0.03), pch=21) +
  geom_tiplab(data=t2$data %>% filter(!`Species Common Name` %in% c("Wall Canyon Sucker", "Owens Sucker", "Warner Sucker")),
              aes(label=`Species Common Name`), size=2, align = TRUE) +
  geom_tiplab(data=t2$data %>% filter(`Species Common Name` %in% c("Wall Canyon Sucker", "Owens Sucker", "Warner Sucker")),
              aes(label=Location), size=2, align = TRUE) +
  geom_cladelab(128,"Owens Sucker", offset=0.08, offset.text=0.001) +
  geom_cladelab(147,"Wall Canyon Sucker", offset=0.08, offset.text=0.001) +
  geom_cladelab(157,"Warner Sucker", offset=0.08, offset.text=0.001) +
#  geom_nodelab() +
  xlim(0,.5) +
  geom_treescale()

ggsave("outputs/204/concat-snps-filtered.pdf", width=7, height=8)
```

