---
title: "Figure-4, Table S2"
author: "Mac Campbell"
date: "2023-02-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message=FALSE)
```

```{r}
library(adegenet)
library(tidyverse)
library(snpR)
library(gplots)
library(viridis)
library(ggpubr)
```

## Let's calculates some pop gen stats with pruned genotypes

Have .vcf in outputs/700/plink.vcf

```{sh, eval=FALSE}
bcftools +prune -l 0.6 -w 10000 outputs/700/plink.vcf -Ov -o outputs/701/plink-pruned.vcf
```

This reduces our variant count to 9511
```{r, eval=FALSE}
vcf<-read.vcfR(file="outputs/701/plink-pruned.vcf")
genind<-vcfR2genind(vcf)
genind@pop<-meta$Location
save(genind, file="outputs/701/pruned-genind.rda")
```

Sanity check of signal:

```{r}
load("outputs/701/pruned-genind.rda")
```

Sanity check of signal:    

```{r}
gen<-genind
X <- tab(gen, NA.method="mean")

pca1 <- dudi.pca(X,scannf=FALSE,scale=FALSE,nf=3)
plot(pca1$li)
```

## Import to snpR

```{r}
meta<-read_tsv("outputs/700/wow.tsv")
sample_meta<-read_tsv("outputs/700/sample_meta.tsv")
my.dat <- import.snpR.data("outputs/701/plink-pruned.vcf", sample.meta = sample_meta)
```



```{r}
x<-calc_pairwise_fst(my.dat, "Group")
groupfst<-get.snpR.stats(x, "Group", "fst")
#this is weighted means
mg<-groupfst$fst.matrix$Group
sample_meta %>% group_by(Species, Group) %>% summarize()
as.tibble(mg)
```

```{r}
y<-calc_pairwise_fst(my.dat, "Species")
speciesfst<-get.snpR.stats(y, "Species", "fst")
#this is weighted means
speciesfst$fst.matrix$Species %>% as_tibble()
```

```{r}
z<-calc_pairwise_fst(my.dat, "Location")
locfst<-get.snpR.stats(z, "Location", "fst")
#this is weighted means
m<-locfst$fst.matrix$Location
```

```{r}
m
```

```{r}

makeSymmLower <- function(m) {
   m[lower.tri(m)] <- t(m)[lower.tri(m)]
   return(m)
}
```


#This code doesn't work as there isn't a zero diagonal
```{r}
m2<-makeSymmLower(as.matrix(m)[,2:9])
m2
```

```{r}
md<- m2 %>% as_tibble() 
mdd<-bind_cols(as_tibble(as.matrix(m[,1])), md)
```

```{r}
md2<-mdd %>% pivot_longer(2:9) %>% rename(`Location 1`=p1, `Location 2`=name, Fst=value) 

```

```{r}
md3<-md2 
md3$Fst<-as.numeric(md3$Fst)

#md3$`Location 1`<-factor(md3$`Location 1`, levels=c("HoneyCreek","TwentymileCreek","DeepCreek","LowerHotCreek","LowerRockCreek",
 #                                                   "LowerHortonCreek","HortonCreek", "SouthFork"))
#md3$`Location 2`<-factor(md3$`Location 2`, levels=c("WallCanyon","HoneyCreek","TwentymileCreek","LowerHotCreek","LowerRockCreek",
 #                                                   "LowerHortonCreek","HortonCreek", "SouthFork"))
```

```{r}
ggplot(md3) + 
  geom_raster(aes(x=`Location 1`, y=`Location 2`, fill=Fst)) +
  geom_text(aes(x=`Location 1`, y=`Location 2`, label=round(Fst,2)), color="white") +
  theme(axis.text.x = element_text(angle=45, vjust=0.5))
```

```{r}
mat_num <- matrix(as.numeric(m2),    # Convert to numeric matrix
                  ncol = ncol(m2))
#mat_num[mat_num<0]<-0
colnames(mat_num)<-colnames(m2)
rownames(mat_num)<-m[,1]$p1
mat_num

```



```{r}
heatmap.2(mat_num, trace="none",
          key=TRUE, srtCol=45, margins = c(5,5), col=viridis(15, direction=-1))
```

###  removing small sample sizes

```{r}
sample_meta %>% group_by(Location) %>% summarize(Count=n())
```

```{r}
my.dat2<-my.dat[,my.dat@sample.meta$Location %in% c("DeepCreek","HoneyCreek","LowerHotCreek","LowerRockCreek","TwentymileCreek","WallCanyon")]
```
```{r}
v<-my.dat2@sample.meta$Location
my.dat3<-my.dat2
my.dat3@sample.meta$Plot<-ifelse(v=="WallCanyon","A",ifelse(v=="HoneyCreek","B",ifelse(v=="TwentymileCreek","C",
                         ifelse(v=="DeepCreek","D",ifelse(v=="LowerHotCreek","E",ifelse(v=="LowerRockCreek","F",NA))))))
```



```{r}
a<-calc_pairwise_fst(my.dat3, "Plot")
locfsta<-get.snpR.stats(a, "Plot", "fst")
#this is weighted means
ma<-locfsta$fst.matrix$Plot
ma %>% as_tibble()
```



## What about Theta, Tajima's D?


```{r}
picalc<-calc_pi(my.dat3,"Plot")
pi<-get.snpR.stats(picalc,"Plot","pi")
```

Generate barplots of pi

```{r}
pidf<-pi$single %>% as_tibble() %>%
  mutate(Species=ifelse(subfacet=="A", "Wall Canyon",
                        ifelse(subfacet %in% c("B","C","D"), "Warner", 
                        ifelse(subfacet %in% c("E","F"), "Owens", "NA"))))
pidf
```

ifelse(v=="WallCanyon","A",ifelse(v=="HoneyCreek","B",ifelse(v=="TwentymileCreek","C",
                         ifelse(v=="DeepCreek","D",ifelse(v=="LowerHotCreek","E",ifelse(v=="LowerRockCreek","F",NA))))))
                         
```{r}
piplot<-ggplot(pidf %>% group_by(Species,subfacet,CHROM) %>% summarize(meanpi=mean(pi))) +
  geom_boxplot(aes(x=subfacet,y=meanpi, fill=Species), alpha=0.75) +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  ylab("Pi") +
  scale_x_discrete(labels=c("Wall Canyon", "Honey Creek","Twentymile Creek","Deep Creek", "Lower Hot Creek", "Lower Rock Creek")) +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  scale_fill_manual(values=alpha(c("red","black","blue"),0.75)) +
  xlab("\nSampling Location") +
  theme(legend.position = "")

piplot
```

```{r}
pidf %>% group_by(Species,subfacet,CHROM) %>% summarize(meanpi=mean(pi)) %>% select(-CHROM) %>% summarize(mean(meanpi))
```

```{r}
pi$weighted.means
```

Calculating tajima's per chrom (takes a minute)
```{r, eval=FALSE}
theta<-calc_tajimas_d(my.dat3, c("Plot.CHROM"))
td<-get.snpR.stats(theta,"Plot.CHROM","tajimas_d")
tddf<-td$single.window %>% as_tibble() %>%
  mutate(Species=ifelse(subfacet=="A", "Wall Canyon",
                        ifelse(subfacet %in% c("B","C","D"), "Warner", 
                        ifelse(subfacet %in% c("E","F"), "Owens", "NA"))))
write_tsv(tddf, "outputs/701/tddf.tsv")
```

```{r}
tddf<-read_tsv("outputs/701/tddf.tsv")
```
```{r}
tplot<-ggplot(tddf) +
  geom_boxplot(aes(x=subfacet,y=D, fill=Species), alpha=0.75) +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  ylab("Tajima's D") +
  scale_x_discrete(labels=c("Wall Canyon", "Honey Creek","Twentymile Creek","Deep Creek", "Lower Hot Creek", "Lower Rock Creek")) +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  scale_fill_manual(values=alpha(c("red","black","blue"),0.75)) +
  xlab("\nSampling Location") +
  theme(legend.position = "")

tplot
```

```{r}
tddf %>% group_by(Species, subfacet) %>% mutate(Tajimas=mean(D)) %>% select(Tajimas) %>% unique()
```


```{r}
ggarrange(piplot + ggtitle("A") + theme(plot.title = element_text(size=18, face="bold")), 
          tplot  + ggtitle("B") + theme(plot.title = element_text(size=18, face="bold")))
ggsave("outputs/701/plot.pdf", width=8.5, height = 4)
```


Test if PI are different? there are quite a few options

```{r}
A=pidf %>% filter(subfacet=="A")
B=pidf %>% filter(subfacet=="B")
C=pidf %>% filter(subfacet=="C")
```

Are our data normally distributed?    

```{r}
A
```


```{r}
tab<-wilcox.test(A$pi, B$pi)
```

```{r}
tab
```

Get all options:

```{r}
mydf<-data.frame()
for(i in seq(1:15)) {

#print (i);
vector<-comb[,i]
a<-vector[1]
b<-vector[2]

v1<-pidf %>% filter(subfacet==vector[1])
v2<-pidf %>% filter(subfacet==vector[2])

PiA<-mean(v1$pi)
PiB<-mean(v2$pi)


test1<-wilcox.test(v1$pi, v2$pi)
p.valuePi<-test1$p.value
Species1<-unique(v1$Species)
Species2<-unique(v2$Species)

#t
v1<-tddf %>% filter(subfacet==vector[1])
v2<-tddf %>% filter(subfacet==vector[2])
test2<-wilcox.test(v1$D, v2$D)
p.valueD<-test2$p.value

DA<-mean(v1$D)
DB<-mean(v2$D)

row<-cbind(a,b,Species1, Species2, PiA, PiB, p.valuePi, DA, DB, p.valueD) %>% as_tibble()
#row<-cbind(a,b,Species1, Species2, p.valuePi, p.valueD) %>% as_tibble()

mydf<-bind_rows(mydf, row)
}
```

c("Wall Canyon", "Honey Creek","Twentymile Creek","Deep Creek", "Lower Hot Creek", "Lower Rock Creek")
```{r}
df<-mydf %>% mutate(Group1=ifelse(a=="A", "Wall Canyon", 
                       ifelse(a=="B", "Honey Creek",
                       ifelse(a=="C", "Twentymile Creek",
                       ifelse(a=="D", "Deep Creek",
                       ifelse(a=="E", "Lower Hot Creek",
                       ifelse(a=="F", "Lower Rock Creek",a)))))))
df<-df %>% mutate(Group2=ifelse(b=="A", "Wall Canyon", 
                       ifelse(b=="B", "Honey Creek",
                       ifelse(b=="C", "Twentymile Creek",
                       ifelse(b=="D", "Deep Creek",
                       ifelse(b=="E", "Lower Hot Creek",
                       ifelse(b=="F", "Lower Rock Creek",b)))))))

write_csv(df,"outputs/701/pi-comparisons.csv")
```
